---
title: "Benchmarking analysis of real world data"
author: "Pratibha Panwar"
date: "2025-09-10"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load libraries

```{r}
library(SpatialExperiment)
library(dplyr)
library(peakRAM)
library(anndataR)

library(clustSIGNAL)
library(scran)
library(scater)

library(Banksy)
library(Seurat)
library(harmony)
library(tidyr)
library(data.table)

library(BASS)

library(SpatialPCA)
library(bluster)

library(aricode)

# loading method wrapper functions
source("./scripts/benchmark_funcs.R")
py_file <- "./scripts/benchmarking_methods_py.py"
```




# SeqFISH mouse embryo data

Raw gene expression counts and metadata downloaded from [Spatial Mouse Atlas 2020](https://content.cruk.cam.ac.uk/jmlab/SpatialMouseAtlas2020/), and stored 
in a spe object.

```{r}
spe_me <- readRDS("./spe_mouseEmbryo.Rds") # spe object of downloaded data
```

```{r}
#### data pre-processing ####

# removing low quality cells
spe_me <- spe_me[, spe_me$celltype_mapped_refined != "Low quality"]
spe_me$celltype_mapped_refined <- factor(spe_me$celltype_mapped_refined)
```

```{r}
#### ClustSIGNAL ####

set.seed(567)
mem_me_cSignal <- peakRAM(cSIGNAL_me <- clustSIGNAL(spe_me, samples = "embryo", 
                                                    threads = 1, outputs = "a"))
mem_me_cSignal <- cbind(mem_me_cSignal, Threads = 1, Batch_correct = FALSE,
                        Batch_by = "None", Dataset = "ME", Method = "ClustSIGNAL")

saveRDS(cSIGNAL_me, file = "./data_outputs/benchmarkOut/res_me_cSIGNAL.Rds")
saveRDS(cSIGNAL_me$spe_final, file = "./data_outputs/benchmarkOut/spe_me_cSIGNAL.Rds")
```

```{r}
#### BANKSY ####

mem_me_banksy <- peakRAM(banksy_me <- runBANKSY(
  spe_me, SEED = 567, annots_label = "celltype_mapped_refined", 
  sample_label = "embryo"))
mem_me_banksy <- cbind(mem_me_banksy, Threads = 1, Batch_correct = FALSE,
                       Batch_by = "None", Dataset = "ME", Method = "BANKSY")

saveRDS(banksy_me, file = "./data_outputs/benchmarkOut/spe_me_banksy.Rds")
```

```{r}
#### BASS ####

# parameters:
#   C = 23 cell types (based on manual annotations)
#   R = 10 domains (randomly chosen value)
#   counts list = cntm_me
#   coordinates list = xym_me

sampleNames <- unique(spe_me$embryo)
cntm_me <- lapply(sampleNames, function(x) counts(spe_me[, spe_me$embryo == x]))
xym_me <- lapply(sampleNames, function(x) spatialCoords(spe_me[, spe_me$embryo == x]))

set.seed(567)
mem_me_bass <- peakRAM(bass_me <- runBASS(cntm_me, xym_me, C = 23, R = 10, 
                                          batch = F))
mem_me_bass <- cbind(mem_me_bass, Threads = 1, Batch_correct = FALSE,
                     Batch_by = "None", Dataset = "ME", Method = "BASS")

saveRDS(bass_me, file = "./data_outputs/benchmarkOut/BASSobj_me.Rds")
```

```{r}
#### SpatialPCA ####

set.seed(567)
mem_me_sPCA <- peakRAM(sPCA_me <- runSpatialPCA(spe_me, samples = "embryo"))
mem_me_sPCA <- cbind(mem_me_sPCA, Threads = 1, Batch_correct = FALSE,
                     Batch_by = "None", Dataset = "ME", Method = "SpatialPCA")

saveRDS(sPCA_me, file = "./data_outputs/benchmarkOut/clustList_me_sPCA.Rds")
```

```{r}
#### SpaGCN ####

# loading python environment
py_path <- "/pythonVENVs/SpaGCN_bench/bin/python"
reticulate::use_python(py_path, required = TRUE)

# parameters:
#   n_clusters = 23 cell types (based on manual annotations)

spagcn_res_me <- runSpaGCN(spe = spe_me, sample_label = "embryo", 
                           n_clusters = 23, SEED = 567, py_file = py_file,
                           python_path  = py_path)
saveRDS(spagcn_res_me, file = "./data_outputs/benchmarkOut/res_me_spagcn.Rds")
```

```{r}
#### STAGATE ####

# loading python environment
py_path <- "../../pythonVENVs/STAGATE_pyG_bench/bin/python"
reticulate::use_python(py_path, required = TRUE)


# parameters:
#   n_clusters = 23 cell types (based on manual annotations)

stagate_res_me <- runSTAGATE(spe = spe_me, 
                             sample_label = "embryo", 
                             n_clusters = 23, k_cutoff = 30,
                             SEED = 567, n_epochs = 500,
                             py_file = py_file,
                             python_path  = py_path,
                             cluster_method = "louvain")
saveRDS(stagate_res_me, file = "./data_outputs/benchmarkOut/res_me_stagate.Rds")
```

```{r}
#### GraphST ####

# creating anndata files for GraphST runs
create_annFiles(spe = spe_me, sample_label = "embryo",
                fileSavepath = "./data_outputs/annDataFiles/mouse_embryo/",
                keep_cols = c("uniqueID", "embryo", "pos", "z", "x_global", 
                              "y_global", "celltype_mapped_refined"))

# Method run using the script in benchmarking_GraphST.py
# Cluster labels = "./data_outputs/benchmarkOut/graphst_me.csv"
# Memory and runtime data = "./data_outputs/benchmarkOut/mem_me_graphst.csv"
```

```{r}
#### Combining benchmarking data ####

cSIGNAL_me <- readRDS("./data_outputs/benchmarkOut/spe_me_cSIGNAL.Rds")
banksy_me <- readRDS("./data_outputs/benchmarkOut/spe_me_banksy.Rds")
BASS_me <- readRDS("./data_outputs/benchmarkOut/BASSobj_me.Rds")
sPCA_me <- readRDS("./data_outputs/benchmarkOut/clustList_me_sPCA.Rds")
graphst_me <- read.csv("./data_outputs/benchmarkOut/graphst_me.csv")
spagcn_me <- readRDS("./data_outputs/benchmarkOut/res_me_spagcn.Rds")
spagcn_me_clust <- setNames(unlist(spagcn_me$raw[[1]]$clusters),
                            unlist(spagcn_me$raw[[1]]$barcodes))
stagate_me <- readRDS("./data_outputs/benchmarkOut/res_me_stagate.Rds")
stagate_me_clust <- setNames(unlist(stagate_me$raw[[1]]$clusters),
                             unlist(stagate_me$raw[[1]]$barcodes))


# combining cluster labels obtained from different clustering methods
clustDF_me <- data.frame(Dataset = "ME",
                         Cells = colnames(cSIGNAL_me), 
                         Sample = cSIGNAL_me$embryo,
                         x = spatialCoords(cSIGNAL_me)[, 1],
                         y = spatialCoords(cSIGNAL_me)[, 2],
                         Celltype = cSIGNAL_me$celltype_mapped_refined) %>%
  mutate(ClustSIGNAL = cSIGNAL_me$ClustSIGNAL,
         BANKSY = banksy_me$clust_M1_lam0.2_k50_res1,
         BASS = factor(unlist(BASS_me@results$c)),
         SpatialPCA = factor(unlist(sPCA_me), 
                             levels = c(1:22, "Unlabelled"))[Cells],
         GraphST = 
           factor(graphst_me[order(match(graphst_me$barcode, Cells)),]$louvain),
         # convert any NA labels to "Unlabelled"
         SpaGCN = factor(coalesce(as.character(spagcn_me_clust[Cells]), 
                                  "Unlabelled")),
         STAGATE = factor(coalesce(as.character(stagate_me_clust[Cells]), 
                                   "Unlabelled")))
saveRDS(clustDF_me, file = "./data_outputs/benchmarkOut/BenchDF_me_update.Rds")
```




# Xenium breast cancer data

Raw gene expression counts and metadata downloaded from [10x Genomics website](https://www.10xgenomics.com/products/xenium-in-situ/preview-dataset-human-breast). 
The data were stored as a MoleculeExperiment object, from which an spe object was generated. Cell type annotations were acquired from source data in [Janesick et al, 2023](https://www.nature.com/articles/s41467-023-43458-x).

```{r}
spe_br <- readRDS("./spe_breastCancer.Rds") # spe object of downloaded data
```

```{r}
#### ClustSIGNAL ####

set.seed(567)
mem_br_cSignal <- peakRAM(cSIGNAL_br <- clustSIGNAL(spe_br, samples = "sample_id", 
                                                    threads = 1, outputs = "a"))
mem_br_cSignal <- cbind(mem_br_cSignal, Threads = 1, Batch_correct = FALSE,
                        Batch_by = "None", Dataset = "BR", Method = "ClustSIGNAL")

saveRDS(cSIGNAL_br, file = "./data_outputs/benchmarkOut/res_br_cSIGNAL.Rds")
saveRDS(cSIGNAL_br$spe_final, file = "./data_outputs/benchmarkOut/spe_br_cSIGNAL.Rds")
```

```{r}
#### BANKSY ####

mem_br_banksy <- peakRAM(banksy_br <- runBANKSY(
  spe_br, SEED = 567, annots_label = "celltype", sample_label = "sample_id"))
mem_br_banksy <- cbind(mem_br_banksy, Threads = 1, Batch_correct = FALSE,
                       Batch_by = "None", Dataset = "BR", Method = "BANKSY")

saveRDS(banksy_br, file = "./data_outputs/benchmarkOut/spe_br_banksy.Rds")
```

```{r}
#### BASS ####

# parameters:
#   C = 19 cell types (based on manual annotations)
#   R = 7 domains (based on Fig. 4d in Janesick et al, 2023)
#   counts list = cntm_br
#   coordinates list = xym_br

# NOTE: C stack memory limit needs to be increased to 64MB before running

sampleNames <- unique(spe_br$sample_id)
cntm_br <- lapply(sampleNames, function(x) counts(spe_br[, spe_br$sample_id == x]))
xym_br <- lapply(sampleNames, function(x) spatialCoords(spe_br[, spe_br$sample_id == x]))

set.seed(567)
mem_br_bass <- peakRAM(bass_br <- runBASS(cntm_br, xym_br, C = 19, R = 7, 
                                          batch = F))
mem_br_bass <- cbind(mem_br_bass, Threads = 1, Batch_correct = FALSE,
                     Batch_by = "None", Dataset = "BR", Method = "BASS")

saveRDS(bass_br, file = "./data_outputs/benchmarkOut/BASSobj_br.Rds")
```

```{r}
#### SpatialPCA ####

set.seed(567)
mem_br_sPCA <- peakRAM(sPCA_br <- runSpatialPCA(spe_br, samples = "sample_id"))

# Out of memory: memory spiked to >400 GB
```

```{r}
#### SpaGCN ####

# loading python environment
py_path <- "../../pythonVENVs/SpaGCN_bench/bin/python"
reticulate::use_python(py_path, required = TRUE)

# parameters:
#   C = 19 cell types (based on manual annotations)

spagcn_res_br <- runSpaGCN(spe = spe_br, sample_label = "sample_id", 
                           n_clusters = 19, SEED = 567, py_file = py_file,
                           python_path  = py_path)
saveRDS(spagcn_res_br, file = "./data_outputs/benchmarkOut/res_br_spagcn.Rds")
```

```{r}
#### STAGATE ####

# loading python environment
py_path <- "../../pythonVENVs/STAGATE_pyG_bench/bin/python"
reticulate::use_python(py_path, required = TRUE)

# parameters:
#   C = 19 cell types (based on manual annotations)

stagate_res_br <- runSTAGATE(spe = spe_br, sample_label = "sample_id", 
                             n_clusters = 19, k_cutoff = 30, SEED = 567, 
                             n_epochs = 500, py_file = py_file,
                             python_path = py_path, cluster_method = "louvain")

saveRDS(stagate_res_br, file = "./data_outputs/benchmarkOut/res_br_stagate.Rds")
```

```{r}
#### GraphST ####

# creating anndata files for GraphST runs
create_annFiles(spe = spe_br, sample_label = "sample_id",
                fileSavepath = "./data_outputs/annDataFiles/breast_cancer/",
                keep_cols = c("sample_id", "cell_id", "x_location", 
                              "y_location", "celltype"))

# Method run using the script in benchmarking_GraphST.py

# Out of memory: memory spiked to >400 GB
```

```{r}
#### Combining benchmarking data ####

cSIGNAL_br <- readRDS("./data_outputs/benchmarkOut/spe_br_cSIGNAL.Rds")
banksy_br <- readRDS("./data_outputs/benchmarkOut/spe_br_banksy.Rds")
BASS_br <- readRDS("./data_outputs/benchmarkOut/BASSobj_br.Rds")
spagcn_br <- readRDS("./data_outputs/benchmarkOut/res_br_spagcn.Rds")
spagcn_br_clust <- setNames(unlist(spagcn_br$raw[[1]]$clusters),
                            unlist(spagcn_br$raw[[1]]$barcodes))
stagate_br <- readRDS("./data_outputs/benchmarkOut/res_br_stagate.Rds")
stagate_br_clust <- setNames(unlist(stagate_br$raw[[1]]$clusters),
                             unlist(stagate_br$raw[[1]]$barcodes))

# combining cluster labels obtained from different clustering methods
clustDF_br <- data.frame(Dataset = "BR",
                         Cells = colnames(cSIGNAL_br), 
                         Sample = cSIGNAL_br$sample_id,
                         x = spatialCoords(cSIGNAL_br)[, 1],
                         y = spatialCoords(cSIGNAL_br)[, 2],
                         Celltype = cSIGNAL_br$celltype) %>%
  mutate(ClustSIGNAL = cSIGNAL_br$ClustSIGNAL,
         BANKSY = banksy_br$clust_M1_lam0.2_k50_res1,
         BASS = factor(unlist(BASS_br@results$c)),
         SpatialPCA = NA,
         GraphST = NA,
         # convert any NA labels to "Unlabelled"
         SpaGCN = factor(coalesce(as.character(spagcn_br_clust[Cells]), 
                                  "Unlabelled")),
         STAGATE = factor(coalesce(as.character(stagate_br_clust[Cells]), 
                                   "Unlabelled")))

saveRDS(clustDF_br, file = "./data_outputs/benchmarkOut/BenchDF_br_update.Rds")
```



# CosMx lung cancer data 

A giotto object of processed gene expression and metadata downloaded from 
[Nanostring website](https://nanostring.com/products/cosmx-spatial-molecular-imager/ffpe-dataset/nsclc-ffpe-dataset/). The giotto object was converted to spe object.

```{r}
spe_lc <- readRDS("./spe_lungNSCLC.Rds") # spe object of downloaded data
```

```{r}
#### ClustSIGNAL ####

set.seed(567)
mem_lc_cSignal <- peakRAM(cSIGNAL_lc <- clustSIGNAL(
  spe_lc, samples = "Run_Tissue_name", threads = 1, outputs = "a"))
mem_lc_cSignal <- cbind(mem_lc_cSignal, Threads = 1, Batch_correct = FALSE,
                        Batch_by = "None", Dataset = "LC", Method = "ClustSIGNAL")

saveRDS(cSIGNAL_lc, file = "./data_outputs/benchmarkOut/res_lc_cSIGNAL.Rds")
saveRDS(cSIGNAL_lc$spe_final, file = "./data_outputs/benchmarkOut/spe_lc_cSIGNAL.Rds")
```

```{r}
#### BANKSY ####

mem_lc_banksy <- peakRAM(banksy_lc <- runBANKSY(
  spe_lc, SEED = 567, annots_label = "cell_type", 
  sample_label = "Run_Tissue_name"))
mem_lc_banksy <- cbind(mem_lc_banksy, Threads = 1, Batch_correct = FALSE,
                       Batch_by = "None", Dataset = "LC", Method = "BANKSY")

saveRDS(banksy_lc, file = "./data_outputs/benchmarkOut/spe_lc_banksy.Rds")
```

```{r}
#### BASS ####

# parameters:
#   C = 22 cell types (based on manual annotations)
#   R = 9 domains (based on niches in dataset)
#   counts list = cntm_lc
#   coordinates list = xym_lc

# NOTE: C stack memory limit needs to be increased to 64MB before running

sampleNames <- unique(spe_lc$Run_Tissue_name)
cntm_lc <- lapply(sampleNames, 
                  function(x) counts(spe_lc[, spe_lc$Run_Tissue_name == x]))
xym_lc <- lapply(sampleNames, 
                 function(x) spatialCoords(spe_lc[, spe_lc$Run_Tissue_name == x]))

set.seed(567)
mem_lc_bass <- peakRAM(bass_lc <- runBASS(cntm_lc, xym_lc, C = 22, R = 9,
                                          batch = F))
mem_lc_bass <- cbind(mem_lc_bass, Threads = 1, Batch_correct = FALSE,
                     Batch_by = "None", Dataset = "LC", Method = "BASS")

saveRDS(bass_lc, file = "./data_outputs/benchmarkOut/BASSobj_lc.Rds")
```

```{r}
#### SpatialPCA ####

set.seed(567)
mem_lc_sPCA <- peakRAM(sPCA_lc <- runSpatialPCA(spe_lc, samples = "Run_Tissue_name"))

# Out of memory: memory spiked to >400 GB
```

```{r}
#### SpaGCN ####

# loading python environment
py_path <- "../../pythonVENVs/SpaGCN_bench/bin/python"
reticulate::use_python(py_path, required = TRUE)

# parameters:
#   C = 22 cell types (based on manual annotations)
#   l_search_start = 0.001 (reduced from default 0.01 to run the method)

spagcn_res_lc <- runSpaGCN(spe = spe_lc, sample_label = "Run_Tissue_name", 
                           n_clusters = 22, SEED = 567, l_search_start = 0.001,
                           py_file = py_file, python_path  = py_path)

saveRDS(spagcn_res_lc, file = "./data_outputs/benchmarkOut/res_lc_spagcn.Rds")
```

```{r}
#### STAGATE ####

# loading python environment
py_path <- "../../pythonVENVs/STAGATE_pyG_bench/bin/python"
reticulate::use_python(py_path, required = TRUE)

# parameters:
#   C = 22 cell types (based on manual annotations)

stagate_res_lc <- runSTAGATE(spe = spe_lc, sample_label = "Run_Tissue_name", 
                             n_clusters = 22, k_cutoff = 30, SEED = 567, 
                             n_epochs = 500, py_file = py_file, 
                             python_path  = py_path, cluster_method = "louvain")

saveRDS(stagate_res_lc, file = "./data_outputs/benchmarkOut/res_lc_stagate.Rds")
```

```{r}
#### GraphST ####

# creating anndata files for GraphST runs
create_annFiles(spe = spe_lc, sample_label = "Run_Tissue_name",
                fileSavepath = "./data_outputs/annDataFiles/lung_cancer/",
                keep_cols = c("cell_ID", "fov", "Slide_name", "tissue", 
                              "Run_name", "Run_Tissue_name", "patient", 
                              "cell_type", "niche", "prop_tumor_in_100_neighbors"))

# Method run using the script in benchmarking_GraphST.py

# Out of memory: memory spiked to >400 GB
```

```{r}
#### Combining benchmarking data ####

cSIGNAL_lc <- readRDS("./data_outputs/benchmarkOut/spe_lc_cSIGNAL.Rds")
banksy_lc <- readRDS("./data_outputs/benchmarkOut/spe_lc_banksy.Rds")
BASS_lc <- readRDS("./data_outputs/benchmarkOut/BASSobj_lc.Rds")
spagcn_lc <- readRDS("./data_outputs/benchmarkOut/res_lc_spagcn.Rds")
spagcn_lc_clust <- setNames(unlist(spagcn_lc$raw[[1]]$clusters),
                            unlist(spagcn_lc$raw[[1]]$barcodes))
stagate_lc <- readRDS("./data_outputs/benchmarkOut/res_lc_stagate.Rds")
stagate_lc_clust <- setNames(unlist(stagate_lc$raw[[1]]$clusters),
                             unlist(stagate_lc$raw[[1]]$barcodes))

# combining cluster labels obtained from different clustering methods
clustDF_lc <- data.frame(Dataset = "LC",
                         Cells = colnames(cSIGNAL_lc), 
                         Sample = cSIGNAL_lc$Run_Tissue_name,
                         x = spatialCoords(cSIGNAL_lc)[, 1],
                         y = spatialCoords(cSIGNAL_lc)[, 2],
                         Celltype = factor(cSIGNAL_lc$cell_type)) %>%
  mutate(ClustSIGNAL = cSIGNAL_lc$ClustSIGNAL,
         BANKSY = banksy_lc$clust_PCA_M1_lam0.2_k50_res1,
         BASS = factor(unlist(BASS_lc@results$c)),
         SpatialPCA = NA,
         GraphST = NA,
         # convert any NA labels to "Unlabelled"
         SpaGCN = factor(coalesce(as.character(spagcn_lc_clust[Cells]), 
                                  "Unlabelled")),
         STAGATE = factor(coalesce(as.character(stagate_lc_clust[Cells]), 
                                   "Unlabelled")))

saveRDS(clustDF_lc, file = "./data_outputs/benchmarkOut/BenchDF_lc_update.Rds")
```




# MERFISH mouse hypothalamus

Raw gene expression counts and metadata downloaded from [DRYAD website](https://datadryad.org/dataset/doi:10.5061/dryad.8t8s248), and stored 
in a spe object. Only the 135 genes that were profiled on MERFISH were kept.

```{r}
spe_mh <- readRDS("./spe_mouseHypothal.Rds") # spe object of downloaded data
```

```{r}
#### data pre-processing ####

# removing ambiguous cells - possible doublets
spe_mh <- spe_mh[, spe_mh$Cell_class != "Ambiguous"]
spe_mh$Cell_class <- factor(spe_mh$Cell_class)
```

```{r}
#### ClustSIGNAL ####

set.seed(567)
mem_mh_cSignal <- peakRAM(cSIGNAL_mh <- clustSIGNAL(spe_mh, samples = "samples", 
                                                    threads = 1, outputs = "a"))
mem_mh_cSignal <- cbind(mem_mh_cSignal, Threads = 1, Batch_correct = FALSE,
                        Batch_by = "None", Dataset = "MH", Method = "ClustSIGNAL")

saveRDS(cSIGNAL_mh, file = "./data_outputs/benchmarkOut/res_mh_cSIGNAL.Rds")
saveRDS(cSIGNAL_mh$spe_final, file = "./data_outputs/benchmarkOut/spe_mh_cSIGNAL.Rds")
```

```{r}
#### BANKSY ####

mem_mh_banksy <- peakRAM(banksy_mh <- runBANKSY(
  spe_mh, SEED = 567, annots_label = "Cell_class", sample_label = "samples"))
mem_mh_banksy <- cbind(mem_mh_banksy, Threads = 1, Batch_correct = FALSE,
                       Batch_by = "None", Dataset = "MH", Method = "BANKSY")

saveRDS(banksy_mh, file = "./data_outputs/benchmarkOut/spe_mh_banksy.Rds")
```

```{r}
#### BASS ####

# parameters:
#   C = 16 cell types (based on manual annotations)
#   R = 9 domains (based on Fig. 3e in Moffitt et al, 2018)
#   counts list = cntm_mh
#   coordinates list = xym_mh

sampleNames <- unique(spe_mh$samples)
cntm_mh <- lapply(sampleNames, 
                  function(x) counts(spe_mh[, spe_mh$samples == x]))
xym_mh <- lapply(sampleNames, 
                 function(x) spatialCoords(spe_mh[, spe_mh$samples == x]))

set.seed(567)
mem_mh_bass <- peakRAM(bass_mh <- runBASS(cntm_mh, xym_mh, C = 16, R = 9))
mem_mh_bass <- cbind(mem_mh_bass, Threads = 1, Batch_correct = FALSE,
                     Batch_by = "None", Dataset = "MH", Method = "BASS")

saveRDS(bass_mh, file = "./data_outputs/benchmarkOut/BASSobj_mh.Rds")
```

```{r}
#### SpatialPCA ####

set.seed(567)
mem_mh_sPCA <- peakRAM(sPCA_mh <- runSpatialPCA(spe_mh, samples = "samples"))
mem_mh_sPCA <- cbind(mem_mh_sPCA, Threads = 1, Batch_correct = FALSE,
                     Batch_by = "None", Dataset = "MH", Method = "SpatialPCA")

saveRDS(sPCA_mh, file = "./data_outputs/benchmarkOut/clustList_mh_sPCA.Rds")
```

```{r}
#### SpaGCN ####

# loading python environment
py_path <- "../../pythonVENVs/SpaGCN_bench/bin/python"
reticulate::use_python(py_path, required = TRUE)

# parameters:
#   C = 16 cell types (based on manual annotations)

spagcn_res_mh <- runSpaGCN(spe = spe_mh, sample_label = "samples", 
                           n_clusters = 16, SEED = 567, py_file = py_file,
                           python_path = py_path)

saveRDS(spagcn_res_mh, file = "./data_outputs/benchmarkOut/res_mh_spagcn.Rds")
```

```{r}
#### STAGATE ####

# loading python environment
py_path <- "../../pythonVENVs/STAGATE_pyG_bench/bin/python"
reticulate::use_python(py_path, required = TRUE)

# parameters:
#   C = 16 cell types (based on manual annotations)

stagate_res_mh <- runSTAGATE(spe = spe_mh, sample_label = "samples", 
                             n_clusters = 16, k_cutoff = 30, SEED = 567, 
                             n_epochs = 500, py_file = py_file, 
                             python_path = py_path, cluster_method = "louvain")

saveRDS(stagate_res_mh, file = "./data_outputs/benchmarkOut/res_mh_stagate.Rds")
```

```{r}
#### GraphST ####

# creating anndata files for GraphST runs
create_annFiles(spe = spe_mh, sample_label = "samples",
                fileSavepath = "./data_outputs/annDataFiles/mouse_brain/",
                keep_cols = c("Cell_ID", "Animal_ID", "Animal_sex", "Behavior",
                              "Bregma", "Cell_class", "Neuron_cluster_ID", 
                              "samples" ))

# Method run using the script in benchmarking_GraphST.py

# Exceeded the maximum allowed runtime: runtime >4 days
```

```{r}
#### Combining benchmarking data ####
library(SpatialExperiment)
library(dplyr)

cSIGNAL_mh <- readRDS("./data_outputs/benchmarkOut/spe_mh_cSIGNAL.Rds")
banksy_mh <- readRDS("./data_outputs/benchmarkOut/spe_mh_banksy.Rds")
BASS_mh <- readRDS("./data_outputs/benchmarkOut/BASSobj_mh.Rds")
sPCA_mh <- readRDS("./data_outputs/benchmarkOut/clustList_mh_sPCA.Rds")
spagcn_mh <- readRDS("./data_outputs/benchmarkOut/res_mh_spagcn.Rds")
spagcn_mh_clust <- setNames(unlist(spagcn_mh$raw[[1]]$clusters),
                            unlist(spagcn_mh$raw[[1]]$barcodes))
stagate_mh <- readRDS("./data_outputs/benchmarkOut/res_mh_stagate.Rds")
stagate_mh_clust <- setNames(unlist(stagate_mh$raw[[1]]$clusters),
                             unlist(stagate_mh$raw[[1]]$barcodes))

# combining cluster labels obtained from different clustering methods
clustDF_mh <- data.frame(Dataset = "MH",
                         Cells = colnames(cSIGNAL_mh), 
                         Sample = cSIGNAL_mh$samples,
                         x = spatialCoords(cSIGNAL_mh)[, 1],
                         y = spatialCoords(cSIGNAL_mh)[, 2],
                         Celltype = cSIGNAL_mh$Cell_class) %>%
  mutate(ClustSIGNAL = cSIGNAL_mh$ClustSIGNAL,
         BANKSY = banksy_mh$clust_M1_lam0.2_k50_res1,
         BASS = factor(unlist(BASS_mh@results$c)),
         SpatialPCA = factor(unlist(sPCA_mh), 
                             levels = c(1:12, "Unlabelled"))[Cells],
         GraphST = NA,
         # convert any NA labels to "Unlabelled"
         SpaGCN = factor(coalesce(as.character(spagcn_mh_clust[Cells]), 
                                  "Unlabelled")),
         STAGATE = factor(coalesce(as.character(stagate_mh_clust[Cells]), 
                                   "Unlabelled")))

saveRDS(clustDF_mh, file = "./data_outputs/benchmarkOut/BenchDF_mh_update.Rds")
```




# Stereo-seq human ovarian cancer

Gene expression and metadata were downloaded from https://spatch.pku-genomics.org/#/dataset/stereo, and stored in a spe object.

```{r}
spe_oc <- readRDS("./spe_ovarianCancer.Rds") # spe object of downloaded data
```

```{r}
#### data pre-processing ####

# remove unlabeled cell types
spe_oc <- spe_oc[, !(is.na(spe_oc$annotation))]
# keeping only transcriptome data
spe_oc <- spe_oc[, spe_oc$codex_common == 1]
# removing genes with no expression in any cells
zero_genes <- rowSums(counts(spe_oc)) == 0
spe_oc <- spe_oc[!zero_genes, ]
# converting expression dgRMatrix to dgCMatrix format
col_counts <- as(as(counts(spe_oc), "TsparseMatrix"), "CsparseMatrix")
assay(spe_oc, "counts") <- col_counts
```

```{r}
#### ClustSIGNAL ####
set.seed(567)
mem_cSignal <- peakRAM(cSIGNAL <- runClustSIGNAL(spe_oc, samples = "sample_id", 
                                                 threads = 1, outputs = "a", 
                                                 hvgs = T))
mem_cSignal <- cbind(mem_cSignal, Threads = 1, Batch_correct = FALSE,
                     Batch_by = "None", Dataset = "OC", Method = "ClustSIGNAL")
saveRDS(mem_cSignal, file = "./data_outputs/benchmarkOut/mem_oc_cSignal.Rds")
saveRDS(cSIGNAL$spe_final, file = "./data_outputs/benchmarkOut/spe_oc_cSIGNAL.Rds")
```

```{r}
#### BANKSY ####

mem_banksy <- peakRAM(banksy <- runBANKSY(spe_oc, SEED = 567, 
                                          annots_label = "annotation",
                                          sample_label = "sample_id", hvgs = T))
mem_banksy <- cbind(mem_banksy, Threads = 1, Batch_correct = FALSE,
                    Batch_by = "None", Dataset = "OC", Method = "Banksy")
saveRDS(mem_banksy, file = "./data_outputs/benchmarkOut/mem_oc_banksy.Rds")
saveRDS(banksy, file = "./data_outputs/benchmarkOut/spe_oc_banksy.Rds")
```

```{r}
#### BASS ####

# parameters:
#   C = 13 cell types (based on manual annotations)
#   R = 5 domains (based on manual annotations)
#   counts list - cntm_oc
#   coordinates list - xym_oc

# NOTE: C stack memory limit needs to be increased to 64MB before running

sampleNames <- unique(spe_oc$sample_id)
cntm_oc <- lapply(sampleNames, 
                  function(x) counts(spe_oc[, spe_oc$sample_id == x]))
xym_oc <- lapply(sampleNames, 
                 function(x) spatialCoords(spe_oc[, spe_oc$sample_id == x]))

set.seed(567)
mem_bass <- peakRAM(bass <- runBASS(cntm_oc, xym_oc, C = 13, R = 5, batch = F))
mem_bass <- cbind(mem_bass, Threads = 1, Batch_correct = FALSE,
                  Batch_by = "None", Dataset = "OC", Method = "BASS")

saveRDS(mem_bass, file = "./data_outputs/benchmarkOut/mem_oc_bass.Rds")
saveRDS(bass, file = "./data_outputs/benchmarkOut/BASSobj_oc.Rds")
```

```{r}
#### SpatialPCA ####

set.seed(567)
mem_sPCA <- peakRAM(sPCA <- runSpatialPCA(spe_oc, samples = "sample_id"))

# Out of memory: memory spiked to >400 GB
```

```{r}
#### SpaGCN ####

# loading python environment
py_path <- "../../pythonVENVs/SpaGCN_bench/bin/python"
reticulate::use_python(py_path, required = TRUE)

# parameters:
#   C = 13 cell types (based on manual annotations)

spagcn_res_oc <- runSpaGCN(spe = spe_oc, sample_label = "sample_id", 
                           n_clusters = 13, SEED = 567, py_file = py_file,
                           python_path  = py_path)

# Out of memory: memory spiked to >400 GB
```

```{r}
#### STAGATE ####

# loading python environment
py_path <- "../../pythonVENVs/STAGATE_pyG_bench/bin/python"
reticulate::use_python(py_path, required = TRUE)

# parameters:
#   C = 13 cell types (based on manual annotations)

stagate_res_oc <- runSTAGATE(spe = spe_oc, sample_label = "sample_id", 
                             n_clusters = 13, k_cutoff = 30, SEED = 567, 
                             n_epochs = 500, py_file = py_file,
                             python_path  = py_path, cluster_method = "louvain")

saveRDS(stagate_res_oc, file = "./data_outputs/benchmarkOut/res_oc_stagate.Rds")
```

```{r}
#### GraphST ####

# creating anndata files for GraphST runs
create_annFiles(spe = spe_oc, sample_label = "sample_id",
                fileSavepath = "./data_outputs/annDataFiles/ovarian_cancer/",
                keep_cols = c("annotation", "spatial_cluster", "sample_id"))

# Method run using the script in benchmarking_GraphST.py

# Out of memory: memory spiked to >400 GB
```

```{r}
#### Combining benchmarking data ####

cSIGNAL_oc <- readRDS("./data_outputs/benchmarkOut/spe_oc_cSIGNAL.Rds")
banksy_oc <- readRDS("./data_outputs/benchmarkOut/spe_oc_banksy.Rds")
BASS_oc <- readRDS("./data_outputs/benchmarkOut/BASSobj_oc.Rds")
stagate_oc <- readRDS("./data_outputs/benchmarkOut/res_oc_stagate.Rds")
stagate_oc_clust <- setNames(unlist(stagate_oc$raw[[1]]$clusters),
                             unlist(stagate_oc$raw[[1]]$barcodes))

# combining cluster labels obtained from different clustering methods
clustDF_oc <- data.frame(Dataset = "OC",
                         Cells = colnames(cSIGNAL_oc), 
                         Sample = cSIGNAL_oc$sample_id,
                         x = spatialCoords(cSIGNAL_oc)[, 1],
                         y = spatialCoords(cSIGNAL_oc)[, 2],
                         Celltype = cSIGNAL_oc$spatial_cluster) %>%
  mutate(ClustSIGNAL = cSIGNAL_oc$ClustSIGNAL,
         BANKSY = banksy_oc$clust_PCA_M1_lam0.2_k50_res1,
         BASS = factor(unlist(BASS_oc@results$c)),
         SpatialPCA = NA,
         GraphST = NA,
         SpaGCN = NA,
         # convert any NA labels to "Unlabelled"
         STAGATE = factor(coalesce(as.character(stagate_oc_clust[Cells]), 
                                   "Unlabelled")))

saveRDS(clustDF_oc, file = "./data_outputs/benchmarkOut/BenchDF_oc.Rds")
```




# VisiumHD human COAD

Gene expression and metadata were downloaded from https://spatch.pku-genomics.org/#/dataset/visium, and stored in a spe object.

```{r}
spe_coad <- readRDS("./spe_colonCancer.Rds") # spe object of downloaded data
```

```{r}
#### data pre-processing ####

# remove unlabeled cell types
spe_coad <- spe_coad[, !(is.na(spe_coad$annotation))]
# keeping only transcriptome data
spe_coad <- spe_coad[, spe_coad$codex_common == 1]
# removing genes with no expression in any cells
zero_genes <- rowSums(counts(spe_coad)) == 0
spe_coad <- spe_coad[!zero_genes, ]
# converting expression dgRMatrix to dgCMatrix format
col_counts <- as(as(counts(spe_coad), "TsparseMatrix"), "CsparseMatrix")
assay(spe_coad, "counts") <- col_counts

# adding logcounts
spe_coad <- logNormCounts(spe_coad)
```

```{r}
#### ClustSIGNAL ####

set.seed(567)
mem_cSignal <- peakRAM(cSIGNAL <- runClustSIGNAL(spe_coad, samples = "sample_id", 
                                                 threads = 1, outputs = "a", 
                                                 hvgs = T))
mem_cSignal <- cbind(mem_cSignal, Threads = 1, Batch_correct = FALSE,
                     Batch_by = "None", Dataset = "COAD", Method = "ClustSIGNAL")
saveRDS(mem_cSignal, file = "./data_outputs/benchmarkOut/mem_coad_cSignal.Rds")
saveRDS(cSIGNAL$spe_final, file = "./data_outputs/benchmarkOut/spe_coad_cSIGNAL.Rds")
```

```{r}
#### BANKSY ####

mem_banksy <- peakRAM(banksy <- runBANKSY(spe_coad, SEED = 567, 
                                          annots_label = "annotation",
                                          sample_label = "sample_id", hvgs = T))
mem_banksy <- cbind(mem_banksy, Threads = 1, Batch_correct = FALSE,
                    Batch_by = "None", Dataset = "COAD", Method = "Banksy")
saveRDS(mem_banksy, file = "./data_outputs/benchmarkOut/mem_coad_banksy.Rds")
saveRDS(banksy, file = "./data_outputs/benchmarkOut/spe_coad_banksy.Rds")
```

```{r}
#### BASS ####

# parameters:
#   C = 16 cell types (based on manual annotations)
#   R = 8 domains (based on manual annotations)
#   counts list = cntm_coad
#   coordinates list = xym_coad

# NOTE: C stack memory limit needs to be increased to 64MB before running

# Run started on 5 Jul ~3pm - stopping run at ~1 pm 9 Jul
sampleNames <- unique(spe_coad$sample_id)
cntm_coad <- lapply(sampleNames, 
                    function(x) counts(spe_coad[, spe_coad$sample_id == x]))
xym_coad <- lapply(sampleNames, 
                   function(x) spatialCoords(spe_coad[, spe_coad$sample_id == x]))
set.seed(567)
mem_bass <- peakRAM(bass <- runBASS(cntm, xym, C = 16, R = 8))

# Exceeded the maximum allowed runtime: runtime >4 days
```

```{r}
#### SpatialPCA ####

set.seed(567)
mem_sPCA <- peakRAM(sPCA <- runSpatialPCA(spe_coad, samples = "sample_id"))

# Out of memory: memory spiked to >400 GB
```

```{r}
#### SpaGCN #### 

# loading python environment
py_path <- "../../pythonVENVs/SpaGCN_bench/bin/python"
reticulate::use_python(py_path, required = TRUE)

# parameters:
#   C = 16 cell types (based on manual annotations)

spagcn_res_coad <- runSpaGCN(spe = spe_coad, sample_label = "sample_id", 
                             n_clusters = 16, SEED = 567, py_file = py_file,
                             python_path = py_path)

# Out of memory: memory spiked to >400 GB
```

```{r}
#### STAGATE ####

# loading python environment
py_path <- "../../pythonVENVs/STAGATE_pyG_bench/bin/python"
reticulate::use_python(py_path, required = TRUE)

# parameters:
#   C = 16 cell types (based on manual annotations)

stagate_res_coad <- runSTAGATE(spe = spe_coad, sample_label = "sample_id", 
                               n_clusters = 16, k_cutoff = 30, SEED = 567, 
                               n_epochs = 500, py_file = py_file,
                               python_path = py_path, cluster_method = "louvain")

saveRDS(stagate_res_coad, file = "./data_outputs/benchmarkOut/res_coad_stagate.Rds")
```

```{r}
#### GraphST ####

# creating anndata files for GraphST runs
create_annFiles(spe = spe_coad, sample_label = "sample_id",
                fileSavepath = "./data_outputs/annDataFiles/colon_cancer/",
                keep_cols = c("annotation", "spatial_cluster", "sample_id"))

# Method run using the script in benchmarking_GraphST.py

# Out of memory: memory spiked to >400 GB
```

```{r}
#### Combining benchmarking data ####

cSIGNAL_coad <- readRDS("./data_outputs/benchmarkOut/spe_coad_cSIGNAL.Rds")
banksy_coad <- readRDS("./data_outputs/benchmarkOut/spe_coad_banksy.Rds")
stagate_coad <- readRDS("./data_outputs/benchmarkOut/res_coad_stagate.Rds")
stagate_coad_clust <- setNames(unlist(stagate_coad$raw[[1]]$clusters),
                               unlist(stagate_coad$raw[[1]]$barcodes))

# combining cluster labels obtained from different clustering methods
clustDF_coad <- data.frame(Dataset = "COAD",
                         Cells = colnames(cSIGNAL_coad), 
                         Sample = cSIGNAL_coad$sample_id,
                         x = spatialCoords(cSIGNAL_coad)[, 1],
                         y = spatialCoords(cSIGNAL_coad)[, 2],
                         Celltype = cSIGNAL_coad$spatial_cluster) %>%
  mutate(ClustSIGNAL = cSIGNAL_coad$ClustSIGNAL,
         BANKSY = banksy_coad$clust_PCA_M1_lam0.2_k50_res1,
         BASS = NA,
         SpatialPCA = NA,
         GraphST = NA,
         SpaGCN = NA,
         # convert any NA labels to "Unlabelled"
         STAGATE = factor(coalesce(as.character(stagate_coad_clust[Cells]), 
                                   "Unlabelled")))

saveRDS(clustDF_coad, file = "./data_outputs/benchmarkOut/BenchDF_coad.Rds")
```




# Clustering metrics

```{r}
# loading data frame containing cluster labels and other metadata
clustDF_me <- readRDS("./data_outputs/benchmarkOut/BenchDF_me_update.Rds")
clustDF_br <- readRDS("./data_outputs/benchmarkOut/BenchDF_br_update.Rds")
clustDF_lc <- readRDS("./data_outputs/benchmarkOut/BenchDF_lc_update.Rds")
clustDF_mh <- readRDS("./data_outputs/benchmarkOut/BenchDF_mh_update.Rds")
clustDF_oc <- readRDS("./data_outputs/benchmarkOut/BenchDF_oc.Rds")
clustDF_coad <- readRDS("./data_outputs/benchmarkOut/BenchDF_coad.Rds")

data_list <- list(clustDF_me, clustDF_br, clustDF_lc, clustDF_mh, clustDF_oc,
                  clustDF_coad)

metrics_list <- lapply(data_list, function(df) {

  # NAs in cluster label columns indicate the method did not run on the dataset
  # checking for NAs when calculating ARI 
  check_ari <- function(celltype, cluster) {
    if (all(is.na(cluster))) {
      return(NA)
    }
    return(ARI(celltype, cluster))
  }
  
  # checking for NAs when calculating cluster number 
  check_count <- function(cluster) {
    if (all(is.na(cluster))) {
      return(NA)
    }
    return(length(unique(cluster)))
  }
  
  # summarising information
  df %>% 
    group_by(Dataset, Sample) %>% 
    summarise(
      # ARI values
      ClustSIGNAL_ari = check_ari(Celltype, ClustSIGNAL),
      BANKSY_ari = check_ari(Celltype, BANKSY),
      BASS_ari = check_ari(Celltype, BASS),
      SpatialPCA_ari = check_ari(Celltype, SpatialPCA), 
      GraphST_ari = check_ari(Celltype, GraphST),
      SpaGCN_ari = check_ari(Celltype, SpaGCN),
      STAGATE_ari = check_ari(Celltype, STAGATE),
      
      # total prior annotations
      Annotation_clust = length(unique(Celltype)), 
      
      # total number of clusters
      ClustSIGNAL_clust = check_count(ClustSIGNAL),
      BANKSY_clust = check_count(BANKSY),
      BASS_clust = check_count(BASS),
      SpatialPCA_clust = check_count(SpatialPCA),
      GraphST_clust = check_count(GraphST),
      SpaGCN_clust = check_count(SpaGCN),
      STAGATE_clust = check_count(STAGATE),
      
      .groups = "drop") %>% 
    as.data.frame()
})

metrics_out <- do.call(rbind, metrics_list)

saveRDS(metrics_out, "./data_outputs/benchmarkOut/clustering_metrics_update.Rds")
```




# Computational resources

```{r}
#### computational performance metrics ####

# information from ClustSIGNAL, BANKSY, BASS, SpatialPCA runs
all_resources <- rbind(mem_me_banksy, mem_br_banksy, mem_lc_banksy, 
                       mem_mh_banksy, mem_oc_banksy, mem_coad_banksy, 
                       mem_me_bass, mem_br_bass, mem_lc_bass, mem_mh_bass, 
                       mem_oc_bass, 
                       mem_me_cSignal, mem_br_cSignal, mem_lc_cSignal, 
                       mem_mh_cSignal, mem_oc_cSignal, mem_coad_cSignal,
                       mem_me_sPCA, mem_mh_sPCA)
# information from SpaGCN and STAGATE runs
res_me_spagcn <- readRDS("./data_outputs/benchmarkOut/res_me_spagcn.Rds")$raw[[2]]
res_me_stagate <- readRDS("./data_outputs/benchmarkOut/res_me_stagate.Rds")$raw[[2]]
res_br_spagcn <- readRDS("./data_outputs/benchmarkOut/res_br_spagcn.Rds")$raw[[2]]
res_br_stagate <- readRDS("./data_outputs/benchmarkOut/res_br_stagate.Rds")$raw[[2]]
res_lc_spagcn <- readRDS("./data_outputs/benchmarkOut/res_lc_spagcn.Rds")$raw[[2]]
res_lc_stagate <- readRDS("./data_outputs/benchmarkOut/res_lc_stagate.Rds")$raw[[2]]
res_mh_spagcn <- readRDS("./data_outputs/benchmarkOut/res_mh_spagcn.Rds")$raw[[2]]
res_mh_stagate <- readRDS("./data_outputs/benchmarkOut/res_mh_stagate.Rds")$raw[[2]]
res_oc_stagate <- readRDS("./data_outputs/benchmarkOut/res_oc_stagate.Rds")$raw[[2]]
res_coad_stagate <- readRDS("./data_outputs/benchmarkOut/res_coad_stagate.Rds")$raw[[2]]
# information from GraphST runs
mem_me_graphst <- read.csv("./data_outputs/benchmarkOut/mem_me_graphst.csv")

# combining all resources used
all_resources <- all_resources %>%
  select(Dataset, Method, Elapsed_Time_sec, Peak_RAM_Used_MiB) %>%
  mutate(Method = recode(Method, "Banksy" = "BANKSY")) %>%
  # adding other method outputs
  rbind(
    c("ME", "SpaGCN", (res_me_spagcn$toc - res_me_spagcn$tic), res_me_spagcn$mem),
    c("ME", "STAGATE", (res_me_stagate$toc - res_me_stagate$tic), res_me_stagate$mem),
    c("BR", "SpaGCN", (res_br_spagcn$toc - res_br_spagcn$tic), res_br_spagcn$mem),
    c("BR", "STAGATE", (res_br_stagate$toc - res_br_stagate$tic), res_br_stagate$mem),
    c("LC", "SpaGCN", (res_lc_spagcn$toc - res_lc_spagcn$tic), res_lc_spagcn$mem),
    c("LC", "STAGATE", (res_lc_stagate$toc - res_lc_stagate$tic), res_lc_stagate$mem),
    c("MH", "SpaGCN", (res_mh_spagcn$toc - res_mh_spagcn$tic), res_mh_spagcn$mem),
    c("MH", "STAGATE", (res_mh_stagate$toc - res_mh_stagate$tic), res_mh_stagate$mem),
    c("OC", "STAGATE", (res_oc_stagate$toc - res_oc_stagate$tic), res_oc_stagate$mem),
    c("COAD", "STAGATE", (res_coad_stagate$toc - res_coad_stagate$tic),
      res_coad_stagate$mem),
    c("ME", "GraphST", mem_me_graphst$runtime_sec, mem_me_graphst$peak_memory_MB),
    c("COAD", "BASS", NA, NA),
    c("BR", "SpatialPCA", NA, NA),
    c("LC", "SpatialPCA", NA, NA),
    c("OC", "SpatialPCA", NA, NA),
    c("COAD", "SpatialPCA", NA, NA),
    c("OC", "SpaGCN", NA, NA),
    c("COAD", "SpaGCN", NA, NA),
    c("BR", "GraphST", NA, NA),
    c("LC", "GraphST", NA, NA),
    c("MH", "GraphST", NA, NA),
    c("OC", "GraphST", NA, NA),
    c("COAD", "GraphST", NA, NA)) %>%
  # adding other relevant information
  mutate(Cells = recode(Dataset, "ME" = 52568, "BR" = 274501, "OC" = 284984, 
                        "COAD" = 520382, "LC" = 771236, "MH" = 874756),
         MeanSampleSize = recode(Dataset, "ME" = 17522.67, "BR" = 137250.5, 
                                 "OC" = 284984, "COAD" = 520382, "LC" = 96404.5, 
                                 "MH" = 4832.906),
         CellNum = recode(Dataset, "ME" = "53k", "BR" = "275k", "OC" = "290k", 
                          "COAD" = "520k", "LC" = "771k", "MH" = "875k"),
         Method = factor(Method, levels = c("ClustSIGNAL", "BANKSY", "BASS",
                                            "SpatialPCA", "SpaGCN", "STAGATE",
                                            "GraphST"))) %>%
  arrange(Dataset, Method)
saveRDS(all_resources, "./data_outputs/benchmarkOut/resources_revised.Rds")
```

