---
title: "ClustSIGNAL identifies cell types and subtypes using neighbourhood information adaptively"
author: "Pratibha Panwar"
date: "2025-09-11"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load libraries

```{r}
library(SpatialExperiment)
library(clustSIGNAL)

library(scran)
library(dplyr)
library(reshape2)

library(scater)
library(ggplot2)
library(circlize)
library(ComplexHeatmap)

library(org.Mm.eg.db)
library(AnnotationDbi)
library(clusterProfiler)

library(slingshot)
```

# Panels a-d - SeqFISH mouse embryo analysis

## ClustSIGNAL run without Xist gene

Assessed impact of Xist gene (X-chromosome gene) on clustering, especially 
clustering of brain region. The presence/absence of Xist gene did not affect 
brain region clustering - embryo 2 still showed a distinct forebrain region 
compared to embryos 1 and 3.

```{r}
spe_me <- readRDS("./spe_mouseEmbryo.Rds") # spe object of downloaded data

# removing low quality cells and Xist gene
spe_me <- spe_me[rownames(spe_me) != "Xist", 
                 spe_me$celltype_mapped_refined != "Low quality"]

# running ClustSIGNAL
set.seed(567)
cSIGNAL_me <- clustSIGNAL(spe_me, samples = "embryo", threads = 1, outputs = "s")
spe_me <- cSIGNAL_me$spe_final

# cluster composition of Forebrain/Midbrain/Hindbrain in each sample
table(spe_me$celltype_mapped_refined, 
      spe_me$ClustSIGNAL, spe_me$embryo)["Forebrain/Midbrain/Hindbrain",,]

# spatial plots of ClustSIGNAL clusters without Xist gene
plotReducedDim(spe_me, dimred = "spatial", point_size = 2,
               colour_by = "ClustSIGNAL", scattermore = TRUE) +
  facet_wrap(~spe_me$embryo) +
  scale_color_manual(values = colors)
```

## Forebrain/midbrain/hindbrain cluster analysis

Dataset used for this analysis - mouse embryo data with ClustSIGNAL benchmarking
output.

```{r}
# reading spe object with ClustSIGNAL clusters
spe_me <- readRDS("./data_outputs/benchmarkOut/spe_me_cSIGNAL.Rds")

# clusters associated with Forebrain/Midbrain/Hindbrain region
brain_clusters <- c(10, 13, 12, 19)

# colors
brain_colors <- setNames(c("#B51D8D", "#0F4A9C", "#65A83E", "#cc7818", "grey95"), 
                         c(brain_clusters, "Others"))
```

#### Panel a

```{r}
#### Brain region spatial plots

# spatial plots highlighting Forebrain/Midbrain/Hindbrain region
annotToplot <- ifelse(
  spe_me$celltype_mapped_refined == "Forebrain/Midbrain/Hindbrain",
  "Forebrain/Midbrain/Hindbrain", "Others")
brain_annot <- plotReducedDim(spe_me, dimred = "spatial", 
                              color_by = I(annotToplot), point_size = 3, 
                              point_alpha = 1, scattermore = TRUE) +
  scale_color_manual(values = c("Others" = "grey95", 
                                "Forebrain/Midbrain/Hindbrain" = "#139992")) +
  facet_wrap(~spe_me$embryo, scales = "free") +
  guides(colour = guide_legend(override.aes = list(size = 5))) +
  labs(color = "Annotation") +
  theme_classic() +
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        strip.text = element_blank(),
        strip.background = element_blank(),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 14))

# spatial plots highlighting Forebrain/Midbrain/Hindbrain clusters
clustToplot <- ifelse(spe_me$ClustSIGNAL %in% brain_clusters, 
                      spe_me$ClustSIGNAL, "Others")
brain_clust <- plotReducedDim(
  spe_me, dimred = "spatial", color_by = I(clustToplot), point_size = 3, 
  point_alpha = 1, scattermore = TRUE) +
  facet_wrap(~spe_me$embryo, scales = "free") +
  scale_color_manual(values = brain_colors) +
  guides(colour = guide_legend(override.aes = list(size = 5))) +
  labs(color = "Clusters") +
  theme_classic() +
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        strip.text = element_blank(),
        strip.background = element_blank(),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 14))

# combining spatial plots
comb_brain_plot <- brain_annot / brain_clust + plot_layout(guides = "collect")
```

#### Panel b

```{r}
#### Composition of ClustSIGNAL clusters

# confusion matrix of cell annotations and ClustSIGNAL clusters
cellMatrix <- as.matrix(table(spe_me$celltype_mapped_refined, spe_me$ClustSIGNAL))

# normalising values by number of cells in a cluster
cellMatrix_norm <- sweep(cellMatrix, 2, colSums(cellMatrix), FUN = "/") 

# assigning color range for heatmap
col_func <- colorRamp2(seq(min(cellMatrix_norm),
                           max(cellMatrix_norm), length = 3),
                       c("aliceblue", "#bb0c00", "darkred"))

# ordering clusters
column_ord <- c(10, 13, 12, 19, 4, 14, 22, 20, 3, 21, 17, 7, 9, 16, 11, 5, 6, 
                15, 1, 18, 8, 2) |> as.character()
Heatmap(cellMatrix_norm, name = "Proportion of\ncluster cells", col = col_func,
        border_gp = grid::gpar(col = "black", lty = 2),
        rect_gp = grid::gpar(col = "white", lwd = 2),
        row_title = "Annotations",
        column_title = "Clusters",
        column_title_side = "bottom",
        column_order = column_ord,
        show_row_dend = FALSE,
        show_column_dend = FALSE,
        column_title_gp = gpar(fontsize = 14),
        row_names_gp = gpar(fontsize = 12),
        column_names_gp = gpar(fontsize = 12),
        heatmap_legend_param = list(title_gp = gpar(fontsize = 14),
                                    labels_gp = gpar(fontsize = 12)))
```

#### Panel c

```{r}
#### top markers of brain region clusters

# sub-setting data to Forebrain/Midbrain/Hindbrain clusters and 
# excluding Xist gene from analysis
spe_me_sub <- spe_me[rownames(spe_me) != "Xist", 
                     spe_me$ClustSIGNAL %in% brain_clusters]

# identifying cluster markers
markers <- findMarkers(spe_me_sub, test.type = "t", 
                       groups = as.character(spe_me_sub$ClustSIGNAL))

# plotting top 10 markers from each cluster
top_me_markers <- lapply(markers, function(x) {
  imp_markers <- x[x$FDR < 0.05,] # only consider significant markers
  return(rownames(imp_markers)[1:10])}) |> 
  unlist() |> unique()

# data frame of log normalised expression of top marker genes
gexp_me_marker <- logcounts(spe_me_sub)[top_me_markers, ] |> 
  as.matrix() |> t() |> as.data.frame()
# adding cluster label to expression data frame
gexp_me_marker$cluster <- factor(as.character(spe_me_sub$ClustSIGNAL))
gexp_me_marker_melt <- reshape2::melt(gexp_me_marker, id.vars = "cluster", 
                                      variable.name = "gene", 
                                      value.name = "expression")

# summarizing data - mean gene expression in cluster and 
# percentage of cluster cells in which gene is expressed
summary_data <- gexp_me_marker_melt %>%
  group_by(cluster, gene) %>%
  summarise(
    meanExpression = mean(expression),
    PerctageCells = sum(expression > 0) / n() * 100) %>%
  mutate(cluster = factor(cluster, levels = brain_clusters))

markerPlot_me <- summary_data %>%
  ggplot(aes(x = cluster, y = gene, size = PerctageCells, 
             color = meanExpression)) +
  geom_point() +
  scale_size_continuous(limit = c(0, 100), range = c(0, 5)) + 
  scale_color_gradient(low = "white", high = "darkblue") + 
  labs(x = "Clusters", y = "Genes", color = "Average\nexpression", 
       size = "Percentage cells\nexpressed", 
       title = "Fore/Mid/Hind-brain clusters") +
  theme_classic() +
  theme(axis.title = element_text(size = 14),
        axis.text = element_text(size = 12),
        plot.title = element_text(size = 14),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 12)) 
```

#### Panel d

```{r}
#### Enrichment analysis of the brain clusters

set.seed(567)

# function for assessing and plotting enrichment
enrich_plot <- function(sig_genes, path_num = 10) {
  # sig_genes = names of significant DEGs
  # path_num = number of pathways to plot
  
  go_enrich <- enrichGO(gene = sig_genes,
                        OrgDb = org.Mm.eg.db,
                        keyType = "SYMBOL",
                        ont = "BP",
                        pAdjustMethod = "BH",
                        pvalueCutoff  = 0.05,
                        qvalueCutoff  = 0.05)
  ova_plot <- ggplot(go_enrich[go_enrich$Count >= 5,][1:path_num,],
                     aes(x = reorder(Description, p.adjust, decreasing = T),
                         y = -log10(p.adjust), fill = p.adjust)) +
    geom_bar(stat = "identity") +
    geom_text(aes(label = Count), hjust = 0) +
    coord_flip() +
    scale_fill_gradient(low = "#4c956c", high = "#fefee3",
                        name = "Adjusted\nP-value") +
    labs(y = "-Log10(adjusted P-value)") +
    theme_classic() +
    theme(axis.text = element_text(size = 10),
          axis.title.x = element_text(size = 12),
          axis.title.y = element_blank(),
          legend.title = element_text(size = 12),
          legend.text = element_text(size = 10),
          legend.key.height = unit(0.3, "cm"))
  return(ova_plot)
}

# plotting pathway enrichment
plot_clust10 <- 
  enrich_plot(rownames(markers[["10"]][markers[["10"]]$FDR < 0.05 & 
                                         markers[["10"]]$summary.logFC > 0.2, ]),100)
plot_clust4 <- 
  enrich_plot(rownames(markers[["4"]][markers[["4"]]$FDR < 0.05 & 
                                        markers[["4"]]$summary.logFC > 0.2, ]))
plot_clust12 <- 
  enrich_plot(rownames(markers[["12"]][markers[["12"]]$FDR < 0.05 & 
                                         markers[["12"]]$summary.logFC > 0.2, ]))
plot_clust13 <- 
  enrich_plot(rownames(markers[["13"]][markers[["13"]]$FDR < 0.05 & 
                                         markers[["13"]]$summary.logFC > 0.2, ]))
```

#### Panel e

```{r}
#### expression plots of brain region marker genes

# removing genes that have more global expression
# rearranging genes according to brain clusters they are found in:
brainClust_genes <- c("Sfrp1", "Fgfr3", "Hoxb1", "Irx3", "En1", "Otx2",
                      "Lhx2", "Six3", "Fezf1", "Dlk1")

# sub-setting data to specific brain clusters genes
spe_me_sub <- spe_me[rownames(spe_me) %in% brainClust_genes,]


minCounts <- min(logcounts(spe_me_sub)) # for scale color
maxCounts <- max(logcounts(spe_me_sub)) # for scale color

genes_plotList <- setNames(lapply(brainClust_genes, function(g) {
  g_plt <- plotReducedDim(spe_me_sub, dimred = "spatial", colour_by = g,
                          scattermore = T, point_size = 3) +
    facet_wrap(~spe_me_sub$embryo, scales = "free", ncol = 1) +
    scale_color_continuous(low = "white", high = "darkblue", 
                           limits = c(minCounts, maxCounts)) +
    labs(title = g, color = "Normalised\nexpression") +
    theme_classic() +
    theme(axis.text = element_blank(),
          axis.ticks = element_blank(),
          axis.title = element_blank(),
          axis.line = element_blank(),
          strip.text = element_blank(),
          strip.background = element_blank(),
          plot.title = element_text(size = 14, hjust = 0.5),
          legend.title = element_text(size = 14),
          legend.text = element_text(size = 12))
}), brainClust_genes)

gene_plot <- wrap_plots(genes_plotList, nrow = 1, guides = "collect")
```

#### Panel f

```{r}
#### cluster-based pseudotime trajectory analysis

color_pal <- setNames(c("#635547", "#8EC792", "#9e6762", "#FACB12", "#3F84AA", 
                        "#0F4A9C", "#EF5A9D", "#C594BF", "#ff891c", "#B51D8D", 
                        "#139992", "#65A83E", "#005579", "#8DB5CE", "#C9EBFB", 
                        "#DFCDE4", "#532C8A", "#8870ad", "#cc7818", "#FBBE92", 
                        "#EF4E22", "#f9decf"), 1:22)

# subset top 20 PCs
reducedDim(spe_me, "PCA_20") <- reducedDim(spe_me, "PCA")[, 1:20]

# run slingshot
spe_me_sling <- slingshot(spe_me, clusterLabels = 'ClustSIGNAL', 
                          reducedDim = 'PCA_20')
spe_me_sling_new <- embedCurves(spe_me, "UMAP")

# generate UMAP with trajectory overlays
umap_coords <- reducedDim(spe_me, "UMAP")
cluster_colors <- color_pal[as.character(spe_me$ClustSIGNAL)]
legend_colors <- color_pal[c(1:22)]

plot(umap_coords, col = cluster_colors, pch = 16, cex = 0.4, axes = FALSE, 
                xlab = "", ylab = "")
lines(SlingshotDataSet(spe_me_sling_new), lwd = 0.5, type = 'curves', 
      col = 'black', lty = 1)
legend(x = "topright", inset = c(-0.25, 0), xpd = TRUE, legend = c(1:22), 
       col = legend_colors, pch = 16, 
       pt.cex = 1.2, cex = 0.8, bty = "n")
```

